---
title: "Assignment 8.3"
author: "Laura Hoffmann"
date: "7/24/2020"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2); library(class); library(gmodels)
knitr::opts_knit$set(root.dir = "C:/Users/Laura/Documents/RStudio/dsc520")
```

```{r}
cluster_data <- read.csv("data/clustering-data.csv", header=TRUE)
```

# A

```{r}
cluster.plot <- ggplot(cluster_data, aes(x,y)) + geom_point() + ggtitle("Clustering Data Plot")
cluster.plot 
```

# B

```{r}
set.seed(58)

# K=2
clusters.2 <- kmeans(cluster_data, 2)
cluster_data$k2 <- as.factor(clusters.2$cluster)
clus2.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k2)) + geom_point() + ggtitle("K-Means Clustering K=2")
clus2.plot

# K=3
clusters.3 <- kmeans(cluster_data, 3)
cluster_data$k3 <- as.factor(clusters.3$cluster)
clus3.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k3)) + geom_point() + ggtitle("K-Means Clustering K=3")
clus3.plot

# K=4
clusters.4 <- kmeans(cluster_data, 4)
cluster_data$k4 <- as.factor(clusters.4$cluster)
clus4.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k4)) + geom_point() + ggtitle("K-Means Clustering K=4")
clus4.plot

# K=5
clusters.5 <- kmeans(cluster_data, 5)
cluster_data$k5 <- as.factor(clusters.5$cluster)
clus5.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k5)) + geom_point() + ggtitle("K-Means Clustering K=5")
clus5.plot

# K=6
clusters.6 <- kmeans(cluster_data, 6)
cluster_data$k6 <- as.factor(clusters.6$cluster)
clus6.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k6)) + geom_point() + ggtitle("K-Means Clustering K=6")
clus6.plot

# K=7
clusters.7 <- kmeans(cluster_data, 7)
cluster_data$k7 <- as.factor(clusters.7$cluster)
clus7.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k7)) + geom_point() + ggtitle("K-Means Clustering K=7")
clus7.plot

# K=8
clusters.8 <- kmeans(cluster_data, 8)
cluster_data$k8 <- as.factor(clusters.8$cluster)
clus8.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k8)) + geom_point() + ggtitle("K-Means Clustering K=8")
clus8.plot

# K=9
clusters.9 <- kmeans(cluster_data, 9)
cluster_data$k9 <- as.factor(clusters.9$cluster)
clus9.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k9)) + geom_point() + ggtitle("K-Means Clustering K=9")
clus9.plot

# K=10
clusters.10 <- kmeans(cluster_data, 10)
cluster_data$k10 <- as.factor(clusters.10$cluster)
clus10.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k10)) + geom_point() + ggtitle("K-Means Clustering K=10")
clus10.plot

# K=11
clusters.11 <- kmeans(cluster_data, 11)
cluster_data$k11 <- as.factor(clusters.11$cluster)
clus11.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k11)) + geom_point() + ggtitle("K-Means Clustering K=11")
clus11.plot

# K=12
clusters.12 <- kmeans(cluster_data, 12)
cluster_data$k12 <- as.factor(clusters.12$cluster)
clus12.plot <- ggplot(cluster_data, mapping = aes(x, y, color = k12)) + geom_point()
clus12.plot
```

# C

```{r echo=TRUE}
myDist<-function(p1,p2) sqrt((p1[,1]-p2[,1])^2+(p1[,2]-p2[,2])^2)
```

### K=2 Distance

```{r echo=TRUE}
dist.2.1 <- (myDist(cluster_data[clusters.2$cluster==1,],clusters.2$centers[1,,drop=FALSE]))
dist.2.2 <- (myDist(cluster_data[clusters.2$cluster==2,],clusters.2$centers[2,,drop=FALSE]))
dist.2 <- sum(dist.2.1, dist.2.2)/nrow(cluster_data)
dist.2
```

### K=3 Distance

```{r echo=TRUE}
dist.3.1 <- (myDist(cluster_data[clusters.3$cluster==1,],clusters.3$centers[1,,drop=FALSE]))
dist.3.2 <- (myDist(cluster_data[clusters.3$cluster==2,],clusters.3$centers[2,,drop=FALSE]))
dist.3.3 <- (myDist(cluster_data[clusters.3$cluster==3,],clusters.3$centers[3,,drop=FALSE]))
dist.3 <- sum(dist.3.1, dist.3.2, dist.3.3)/nrow(cluster_data)
dist.3
```

### K=4 Distance

```{r echo=TRUE}
dist.4.1 <- (myDist(cluster_data[clusters.4$cluster==1,],clusters.4$centers[1,,drop=FALSE]))
dist.4.2 <- (myDist(cluster_data[clusters.4$cluster==2,],clusters.4$centers[2,,drop=FALSE]))
dist.4.3 <- (myDist(cluster_data[clusters.4$cluster==3,],clusters.4$centers[3,,drop=FALSE]))
dist.4.4 <- (myDist(cluster_data[clusters.4$cluster==4,],clusters.4$centers[4,,drop=FALSE]))
dist.4 <- sum(dist.4.1, dist.4.2, dist.4.3, dist.4.4)/nrow(cluster_data)
dist.4
```

### K=5 Distance

```{r echo=TRUE}
dist.5.1 <- (myDist(cluster_data[clusters.5$cluster==1,],clusters.5$centers[1,,drop=FALSE]))
dist.5.2 <- (myDist(cluster_data[clusters.5$cluster==2,],clusters.5$centers[2,,drop=FALSE]))
dist.5.3 <- (myDist(cluster_data[clusters.5$cluster==3,],clusters.5$centers[3,,drop=FALSE]))
dist.5.4 <- (myDist(cluster_data[clusters.5$cluster==4,],clusters.5$centers[4,,drop=FALSE]))
dist.5.5 <- (myDist(cluster_data[clusters.5$cluster==5,],clusters.5$centers[5,,drop=FALSE]))
dist.5 <- sum(dist.5.1, dist.5.2, dist.5.3, dist.5.4, dist.5.5)/nrow(cluster_data)
dist.5
```

### K=6 Distance

```{r echo=TRUE}
dist.6.1 <- (myDist(cluster_data[clusters.6$cluster==1,],clusters.6$centers[1,,drop=FALSE]))
dist.6.2 <- (myDist(cluster_data[clusters.6$cluster==2,],clusters.6$centers[2,,drop=FALSE]))
dist.6.3 <- (myDist(cluster_data[clusters.6$cluster==3,],clusters.6$centers[3,,drop=FALSE]))
dist.6.4 <- (myDist(cluster_data[clusters.6$cluster==4,],clusters.6$centers[4,,drop=FALSE]))
dist.6.5 <- (myDist(cluster_data[clusters.6$cluster==5,],clusters.6$centers[5,,drop=FALSE]))
dist.6.6 <- (myDist(cluster_data[clusters.6$cluster==6,],clusters.6$centers[6,,drop=FALSE]))
dist.6 <- sum(dist.6.1, dist.6.2, dist.6.3, dist.6.4, dist.6.5, dist.6.6)/nrow(cluster_data)
dist.6
```

### K=7 Distance

```{r echo=TRUE}
dist.7.1 <- (myDist(cluster_data[clusters.7$cluster==1,],clusters.7$centers[1,,drop=FALSE]))
dist.7.2 <- (myDist(cluster_data[clusters.7$cluster==2,],clusters.7$centers[2,,drop=FALSE]))
dist.7.3 <- (myDist(cluster_data[clusters.7$cluster==3,],clusters.7$centers[3,,drop=FALSE]))
dist.7.4 <- (myDist(cluster_data[clusters.7$cluster==4,],clusters.7$centers[4,,drop=FALSE]))
dist.7.5 <- (myDist(cluster_data[clusters.7$cluster==5,],clusters.7$centers[5,,drop=FALSE]))
dist.7.6 <- (myDist(cluster_data[clusters.7$cluster==6,],clusters.7$centers[6,,drop=FALSE]))
dist.7.7 <- (myDist(cluster_data[clusters.7$cluster==7,],clusters.7$centers[7,,drop=FALSE]))
dist.7 <- sum(dist.7.1, dist.7.2, dist.7.3, dist.7.4, dist.7.5, dist.7.6, dist.7.7)/nrow(cluster_data)
dist.7
```

### K=8 Distance

```{r echo=TRUE}
dist.8.1 <- (myDist(cluster_data[clusters.8$cluster==1,],clusters.8$centers[1,,drop=FALSE]))
dist.8.2 <- (myDist(cluster_data[clusters.8$cluster==2,],clusters.8$centers[2,,drop=FALSE]))
dist.8.3 <- (myDist(cluster_data[clusters.8$cluster==3,],clusters.8$centers[3,,drop=FALSE]))
dist.8.4 <- (myDist(cluster_data[clusters.8$cluster==4,],clusters.8$centers[4,,drop=FALSE]))
dist.8.5 <- (myDist(cluster_data[clusters.8$cluster==5,],clusters.8$centers[5,,drop=FALSE]))
dist.8.6 <- (myDist(cluster_data[clusters.8$cluster==6,],clusters.8$centers[6,,drop=FALSE]))
dist.8.7 <- (myDist(cluster_data[clusters.8$cluster==7,],clusters.8$centers[7,,drop=FALSE]))
dist.8.8 <- (myDist(cluster_data[clusters.8$cluster==8,],clusters.8$centers[8,,drop=FALSE]))
dist.8 <- sum(dist.8.1, dist.8.2, dist.8.3, dist.8.4, dist.8.5, dist.8.6, dist.8.7, dist.8.8)/nrow(cluster_data)
dist.8
```

### K=9 Distance

```{r echo=TRUE}
dist.9.1 <- (myDist(cluster_data[clusters.9$cluster==1,],clusters.9$centers[1,,drop=FALSE]))
dist.9.2 <- (myDist(cluster_data[clusters.9$cluster==2,],clusters.9$centers[2,,drop=FALSE]))
dist.9.3 <- (myDist(cluster_data[clusters.9$cluster==3,],clusters.9$centers[3,,drop=FALSE]))
dist.9.4 <- (myDist(cluster_data[clusters.9$cluster==4,],clusters.9$centers[4,,drop=FALSE]))
dist.9.5 <- (myDist(cluster_data[clusters.9$cluster==5,],clusters.9$centers[5,,drop=FALSE]))
dist.9.6 <- (myDist(cluster_data[clusters.9$cluster==6,],clusters.9$centers[6,,drop=FALSE]))
dist.9.7 <- (myDist(cluster_data[clusters.9$cluster==7,],clusters.9$centers[7,,drop=FALSE]))
dist.9.8 <- (myDist(cluster_data[clusters.9$cluster==8,],clusters.9$centers[8,,drop=FALSE]))
dist.9.9 <- (myDist(cluster_data[clusters.9$cluster==9,],clusters.9$centers[9,,drop=FALSE]))
dist.9 <- sum(dist.9.1, dist.9.2, dist.9.3, dist.9.4, dist.9.5, dist.9.6, dist.9.7, dist.9.8, dist.9.9)/nrow(cluster_data)
dist.9
```

### K=10 Distance

```{r echo=TRUE}
dist.10.1 <- (myDist(cluster_data[clusters.10$cluster==1,],clusters.10$centers[1,,drop=FALSE]))
dist.10.2 <- (myDist(cluster_data[clusters.10$cluster==2,],clusters.10$centers[2,,drop=FALSE]))
dist.10.3 <- (myDist(cluster_data[clusters.10$cluster==3,],clusters.10$centers[3,,drop=FALSE]))
dist.10.4 <- (myDist(cluster_data[clusters.10$cluster==4,],clusters.10$centers[4,,drop=FALSE]))
dist.10.5 <- (myDist(cluster_data[clusters.10$cluster==5,],clusters.10$centers[5,,drop=FALSE]))
dist.10.6 <- (myDist(cluster_data[clusters.10$cluster==6,],clusters.10$centers[6,,drop=FALSE]))
dist.10.7 <- (myDist(cluster_data[clusters.10$cluster==7,],clusters.10$centers[7,,drop=FALSE]))
dist.10.8 <- (myDist(cluster_data[clusters.10$cluster==8,],clusters.10$centers[8,,drop=FALSE]))
dist.10.9 <- (myDist(cluster_data[clusters.10$cluster==9,],clusters.10$centers[9,,drop=FALSE]))
dist.10.10 <- (myDist(cluster_data[clusters.10$cluster==10,],clusters.10$centers[10,,drop=FALSE]))
dist.10 <- sum(dist.10.1, dist.10.2, dist.10.3, dist.10.4, dist.10.5, dist.10.6, dist.10.7, dist.10.8, dist.10.9, dist.10.10)/nrow(cluster_data)
dist.10
```

### K=11 Distance

```{r echo=TRUE}
dist.11.1 <- (myDist(cluster_data[clusters.11$cluster==1,],clusters.11$centers[1,,drop=FALSE]))
dist.11.2 <- (myDist(cluster_data[clusters.11$cluster==2,],clusters.11$centers[2,,drop=FALSE]))
dist.11.3 <- (myDist(cluster_data[clusters.11$cluster==3,],clusters.11$centers[3,,drop=FALSE]))
dist.11.4 <- (myDist(cluster_data[clusters.11$cluster==4,],clusters.11$centers[4,,drop=FALSE]))
dist.11.5 <- (myDist(cluster_data[clusters.11$cluster==5,],clusters.11$centers[5,,drop=FALSE]))
dist.11.6 <- (myDist(cluster_data[clusters.11$cluster==6,],clusters.11$centers[6,,drop=FALSE]))
dist.11.7 <- (myDist(cluster_data[clusters.11$cluster==7,],clusters.11$centers[7,,drop=FALSE]))
dist.11.8 <- (myDist(cluster_data[clusters.11$cluster==8,],clusters.11$centers[8,,drop=FALSE]))
dist.11.9 <- (myDist(cluster_data[clusters.11$cluster==9,],clusters.11$centers[9,,drop=FALSE]))
dist.11.10 <- (myDist(cluster_data[clusters.11$cluster==10,],clusters.11$centers[10,,drop=FALSE]))
dist.11.11 <- (myDist(cluster_data[clusters.11$cluster==11,],clusters.11$centers[11,,drop=FALSE]))
dist.11 <- sum(dist.11.1, dist.11.2, dist.11.3, dist.11.4, dist.11.5, dist.11.6, dist.11.7, dist.11.8, dist.11.9, dist.11.10, dist.11.11)/nrow(cluster_data)
dist.11
```

### K=12 Distance

```{r echo=TRUE}
dist.12.1 <- (myDist(cluster_data[clusters.12$cluster==1,],clusters.12$centers[1,,drop=FALSE]))
dist.12.2 <- (myDist(cluster_data[clusters.12$cluster==2,],clusters.12$centers[2,,drop=FALSE]))
dist.12.3 <- (myDist(cluster_data[clusters.12$cluster==3,],clusters.12$centers[3,,drop=FALSE]))
dist.12.4 <- (myDist(cluster_data[clusters.12$cluster==4,],clusters.12$centers[4,,drop=FALSE]))
dist.12.5 <- (myDist(cluster_data[clusters.12$cluster==5,],clusters.12$centers[5,,drop=FALSE]))
dist.12.6 <- (myDist(cluster_data[clusters.12$cluster==6,],clusters.12$centers[6,,drop=FALSE]))
dist.12.7 <- (myDist(cluster_data[clusters.12$cluster==7,],clusters.12$centers[7,,drop=FALSE]))
dist.12.8 <- (myDist(cluster_data[clusters.12$cluster==8,],clusters.12$centers[8,,drop=FALSE]))
dist.12.9 <- (myDist(cluster_data[clusters.12$cluster==9,],clusters.12$centers[9,,drop=FALSE]))
dist.12.10 <- (myDist(cluster_data[clusters.12$cluster==10,],clusters.12$centers[10,,drop=FALSE]))
dist.12.11 <- (myDist(cluster_data[clusters.12$cluster==11,],clusters.12$centers[11,,drop=FALSE]))
dist.12.12 <- (myDist(cluster_data[clusters.12$cluster==12,],clusters.12$centers[12,,drop=FALSE]))
dist.12 <- sum(dist.12.1, dist.12.2, dist.12.3, dist.12.4, dist.12.5, dist.12.6, dist.12.7, dist.12.8, dist.12.9, dist.12.10, dist.12.11, dist.12.12)/nrow(cluster_data)
dist.12
```

### Plot

```{r}
kvalues <- c(2,3,4,5,6,7,8,9,10,11,12)
distances <- c(dist.2, dist.3, dist.4, dist.5, dist.6, dist.7, dist.8, dist.9, dist.10, dist.11, dist.12)

kdist.df <- data.frame(cbind(kvalues, distances))

kdist.plot <- ggplot(kdist.df, aes(kvalues, distances)) + geom_point() + geom_line() + ggtitle("K-Value Average Cluster Distances")
kdist.plot
```
### Based on the plot above the elbow point looks like it would be at k=9. This is because adding another cluster does not improve, or lower the average distance much. In fact, in this case adding the next cluster increased the average distance. So 9 seems to be the best k value for this data set. 
